# Leaflet

## Getting Started

Leaflet is a [Javascript library](http://leafletjs.com/) that allows users to
create interactive maps that even look good on mobile devices. The wonderful
folks at RStudio created their own Leaflet [R package](https://github.com/rstudio/leaflet)
which allows R users to create their own Leaflet maps without needing to write
any Javascript. Let's get started making some maps!

First you should of course install Leaflet:

```{r, eval=FALSE}
install.packages("leaflet")
```

Once you have leaflet installed you can start making maps! You can store your
map in a variable like any other value, in this exmaple let's store our map in
a variable called `my_map`. You need to use the `addTiles()` function to add
a base layer to your map, and you can use the **pipe operator** (`%>%`) to
sequentially add features to your map. Here's how you can make a very simple,
basic map:

```{r}
library(leaflet)

my_map <- leaflet() %>%
  addTiles()
my_map
```

Whoa, it's the whole world! In the HTML version of this book you should be able
to move the map around and zoom in and out as you please. You've made your first
interactive map! Leaflet uses free and open source data from the
[OpenStreetMap](https://www.openstreetmap.org/) project.

As lovely as the whole world is, you might want to specifiy a particular 
location and zoom level if you're trying to show a particular location on your
map. You can do this with the `setView()` function by specifying the longitude,
latitude, and the zoom level:

```{r}
my_map <- my_map %>%
  setView(lat = 39.2898036, lng = -76.6051842, zoom = 13)
my_map
```

Ah look it's beautiful Baltimore, the
[Greatest City in America](https://flic.kr/p/aQsDaT)! 

## Adding Markers

This is certainly a more
interesting map than we had before, but it's still lacking features. Let's add
a marker to this map using the `addMarkers()` function:

```{r}
my_map <- my_map %>%
  addMarkers(lat=39.2980803, lng=-76.5898801, 
             popup="Jeff Leek's Office")
my_map
```

As you can see you can add a marker just by specifying longitude and latitude.
Optionally you can add some pop up text which appears when you click on the
marker. You don't have to add one marker at a time. To create multiple markers
at once, use a data frame with columns named `lat` and `lng` for latitude and
longitude:

```{r}
suppressWarnings(set.seed(2016-04-25))
df <- data.frame(lat = runif(20, min = 39.2, max = 39.3),
                 lng = runif(20, min = -76.6, max = -76.5))
df %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers()
```

As you can see you can use the pipe operator to pipe a data frame directly into
`leaflet()`.

The default blue "teardrop" icon that Leaflet provides is nice, but you can also
create your own icon with the `makeIcon()` function. You'll need to provide a
few arguments to this function including a link to the URL for your icon, plus
width and height variables. Once you've made an icon object you can use that
icon by specifying the `icon` argument in `addmarkers()`:

```{r}
hopkinsIcon <- makeIcon(
  iconUrl = "http://brand.jhu.edu/content/uploads/2014/06/university.shield.small_.blue_.png",
  iconWidth = 31*215/230, iconHeight = 31,
  iconAnchorX = 31*215/230/2, iconAnchorY = 16
)

hopkinsLatLong <- data.frame(
  lat = c(39.2973166, 39.3288851, 39.2906617, 39.2970681, 39.2824806),
  lng = c(-76.5929798, -76.6206598, -76.5469683, -76.6150537, -76.6016766))

hopkinsLatLong %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers(icon = hopkinsIcon)
```

Now we've nicely mapped out all of the Johns Hopkins campus locations around the
city of Baltimore.
You can customize text pop ups by including any valid HTML code in a vector.
Then you can provide that vector as the `popup` argument in `addMarkers()` like
so:

```{r}
hopkinsSites <- c(
  "<a href='http://www.jhsph.edu/'>East Baltimore Campus</a>",
  "<a href='https://apply.jhu.edu/visit/homewood/'>Homewood Campus</a>",
  "<a href='http://www.hopkinsmedicine.org/johns_hopkins_bayview/'>Bayview Medical Center</a>",
  "<a href='http://www.peabody.jhu.edu/'>Peabody Institute</a>",
  "<a href='http://carey.jhu.edu/'>Carey Business School</a>"
)

hopkinsLatLong %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers(icon = hopkinsIcon, popup = hopkinsSites)
```

Each Johns Hopkins campus location now has a pop up with a link to the website
for its respective campus!



```{r}
df <- data.frame(lat = runif(500, min = 39.25, max = 39.35),
                 lng = runif(500, min = -76.65, max = -76.55))
df %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers(clusterOptions = markerClusterOptions())
```

```{r}
df <- data.frame(lat = runif(20, min = 39.25, max = 39.35),
                 lng = runif(20, min = -76.65, max = -76.55))
df %>% 
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers()
```

## Drawing

```{r}
md_cities <- data.frame(name = c("Baltimore", "Frederick", "Rockville", "Gaithersburg", 
                                 "Bowie", "Hagerstown", "Annapolis", "College Park", "Salisbury", "Laurel"),
                        pop = c(619493, 66169, 62334, 61045, 55232,
                                39890, 38880, 30587, 30484, 25346),
                        lat = c(39.2920592, 39.4143921, 39.0840, 39.1434, 39.0068, 39.6418, 38.9784, 38.9897, 38.3607, 39.0993),
                        lng = c(-76.6077852, -77.4204875, -77.1528, -77.2014, -76.7791, -77.7200, -76.4922, -76.9378, -75.5994, -76.8483))
md_cities %>%
  leaflet() %>%
  addTiles() %>%
  addCircles(weight = 1, radius = sqrt(md_cities$pop) * 30)
```

```{r}
leaflet() %>%
  addTiles() %>%
  addRectangles(lat1 = 37.3858, lng1 = -122.0595, 
                lat2 = 37.3890, lng2 = -122.0625)
```

```{r, eval=FALSE}
df <- data.frame(lat = runif(20, min = 39.25, max = 39.35),
                 lng = runif(20, min = -76.65, max = -76.55),
                 col = sample(c("red", "blue", "green"), 20, replace = TRUE),
                 stringsAsFactors = FALSE)

df %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(color = df$col) %>%
  addLegend(labels = LETTERS[1:3], colors = c("blue", "red", "green"))
```

## Conclusion

For more details about the leaflet package for R 
visit http://rstudio.github.io/leaflet/.